
# Course project report

## Load dataset

```{r}
library(caret)
dataset <- read.csv("./pml-training.csv")
```

## Plot and analyze predictors

```{r}
names(dataset)
summary(dataset)
plot(dataset$roll_arm)
qplot(roll_arm, roll_belt, colour = classe,data = dataset)
```

### Remove near-zero variables, non predictor variables, ….

```{r}
dataselect <- names(dataset[,-c(1:7,12:36,50:59,69:83,87:94,95:98,99:101,103:115,125:139,141:150)])
adataset <- dataset[,dataselect]; names(adataset)
```

Below are selected predictors:

[1] "roll_belt"            "pitch_belt"           "yaw_belt"            
 [4] "total_accel_belt"     "gyros_belt_x"         "gyros_belt_y"        
 [7] "gyros_belt_z"         "accel_belt_x"         "accel_belt_y"        
[10] "accel_belt_z"         "magnet_belt_x"        "magnet_belt_y"       
[13] "magnet_belt_z"        "roll_arm"             "pitch_arm"           
[16] "yaw_arm"              "total_accel_arm"      "gyros_arm_x"         
[19] "gyros_arm_y"          "gyros_arm_z"          "accel_arm_x"         
[22] "accel_arm_y"          "accel_arm_z"          "magnet_arm_x"        
[25] "magnet_arm_y"         "magnet_arm_z"         "roll_dumbbell"       
[28] "pitch_dumbbell"       "yaw_dumbbell"         "total_accel_dumbbell"
[31] "accel_dumbbell_x"     "accel_dumbbell_y"     "accel_dumbbell_z"    
[34] "magnet_dumbbell_x"    "magnet_dumbbell_y"    "magnet_dumbbell_z"   
[37] "roll_forearm"         "pitch_forearm"        "yaw_forearm"         
[40] "total_accel_forearm"  "gyros_forearm_x"      "gyros_forearm_y"     
[43] "gyros_forearm_z"      "accel_forearm_x"      "accel_forearm_y"     
[46] "accel_forearm_z"      "magnet_forearm_x"     "magnet_forearm_y"    
[49] "magnet_forearm_z"     "classe"              


## Fit a random forest model with cross-validation

```{r}
tc <- trainControl(method="repeatedcv", number=10, repeats=10, classProbs = TRUE)
set.seed(13568);  rfModFit <- train(classe ~ ., data= adataset, method="rf", trControl=tc, ntree=10)
```


## Analyze model accuracy

```{r}
rfModFit
confusionMatrix(rfModFit)
```

Random Forest 

19622 samples
   49 predictors
    5 classes: 'A', 'B', 'C', 'D', 'E' 

No pre-processing
Resampling: Cross-Validated (10 fold, repeated 10 times) 

Summary of sample sizes: 17660, 17659, 17661, 17660, 17660, 17659, ... 

Resampling results across tuning parameters:

  mtry  Accuracy  Kappa  Accuracy SD  Kappa SD
  2     0.985     0.981  0.00291      0.00369 
  25    0.991     0.989  0.00234      0.00296 
  49    0.985     0.981  0.00337      0.00426 

Accuracy was used to select the optimal model using  the largest value.
The final value used for the model was mtry = 25. 

Cross-Validated (10 fold, repeated 10 times) Confusion Matrix 

(entries are percentages of table totals)

          Reference
Prediction    A    B    C    D    E
         A 28.4  0.2  0.0  0.0  0.0
         B  0.0 19.1  0.1  0.0  0.0
         C  0.0  0.1 17.2  0.2  0.0
         D  0.0  0.0  0.1 16.1  0.1
         E  0.0  0.0  0.0  0.0 18.3


### OOB estimate

```{r}
rfModFit$finalModel
```

Call:
randomForest(x = x, y = y, ntree = 10, mtry = param$mtry) 
               Type of random forest: classification
                     Number of trees: 10
No. of variables tried at each split: 25

        OOB estimate of  error rate: 2.59%
Confusion matrix:
     A    B    C    D    E class.error
A 5488   18    8    5    6 0.006696833
B   77 3622   40   11   15 0.037981408
C    6   66 3278   38    4 0.033608491
D   12   17   73 3071   12 0.035792779
E    4   38   24   29 3482 0.026558569


### Look at variable importance

```{r}
varImp(rfModFit)
varImpPlot(rfModFit$finalModel)
```

rf variable importance

  only 20 most important variables shown (out of 49)

                     Overall
roll_belt             100.00
pitch_forearm          62.99
pitch_belt             56.50
yaw_belt               54.89
magnet_dumbbell_y      47.80
magnet_dumbbell_z      44.39
roll_forearm           38.69
accel_dumbbell_y       26.47
magnet_dumbbell_x      20.49
accel_forearm_x        18.81
magnet_forearm_z       18.70
accel_belt_z           18.19
roll_dumbbell          16.61
magnet_belt_z          15.03
accel_dumbbell_z       13.56
total_accel_dumbbell   13.51
gyros_belt_z           13.49
roll_arm               12.67
yaw_arm                12.41
magnet_arm_x           11.05


### Compute in band accuracy

```{r}
confusionMatrix(adataset$classe, predict(rfModFit, newdata=adataset))
```

Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 5580    0    0    0    0
         B    2 3794    1    0    0
         C    0    0 3422    0    0
         D    0    0    0 3216    0
         E    0    0    0    0 3607

Overall Statistics
                                     
               Accuracy : 0.9998     
                 95% CI : (0.9996, 1)
    No Information Rate : 0.2845     
    P-Value [Acc > NIR] : < 2.2e-16  
                                     
                  Kappa : 0.9998     
 Mcnemar's Test P-Value : NA         

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            0.9996   1.0000   0.9997   1.0000   1.0000
Specificity            1.0000   0.9998   1.0000   1.0000   1.0000
Pos Pred Value         1.0000   0.9992   1.0000   1.0000   1.0000
Neg Pred Value         0.9999   1.0000   0.9999   1.0000   1.0000
Prevalence             0.2845   0.1934   0.1744   0.1639   0.1838
Detection Rate         0.2844   0.1934   0.1744   0.1639   0.1838
Detection Prevalence   0.2844   0.1935   0.1744   0.1639   0.1838
Balanced Accuracy      0.9998   0.9999   0.9999   1.0000   1.0000


## Predict on new dataset

```{r}
testing <- read.csv("./pml-testing.csv")
testselect <- dataselect[-50]
atesting <- testing[,testselect]; names(atesting)
atesting$problem_id <- testing$problem_id
predict(rfModFit, newdata=atesting[,-50])
```

[1] B A B A A E D B A A B C B A E E A B B B
Levels: A B C D E
